---
title: "World Meteorological Data"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: yes
---

## Introduction

Let's practice the following `dplyr` functions to:

1.  select columns
2.  join data, creating a wide dataset
3.  try to plot wide data
4.  pivot data, to make it long
5.  try to plot long data

### Packages

We will use the R `dplyr` package contained in the **tidyverse** library for data wrangling and `ggplot2` for data visualization (also in **tidyverse**)

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

### Data

The data we're using originally came from the [World Meteorological Organization](https://worldweather.wmo.int/en/home.html).

The data comes in separate `.csv` files called `wmo_[city-name].csv` and has already been added to this exercise `/data` folder.
It contains maximum monthly temperatures in degrees Fahrenheit from long-term climate averages.

The code chunk below will read the different data sets in as objects.
Check your environment pane to make sure they loaded into your RStudio session after you run the chunk.

```{r message = F}
csv_files <- list.files("data/", pattern = ".csv")
data_list <- lapply(paste0("data/",csv_files), read_csv) 

wmo_chicago <- as_tibble(data_list[[1]])
wmo_dhaka <- as_tibble(data_list[[2]])
wmo_fairbanks <- as_tibble(data_list[[3]])
wmo_gabarone <- as_tibble(data_list[[4]])
wmo_kiev <- as_tibble(data_list[[5]])
wmo_perth <- as_tibble(data_list[[6]])
wmo_rio_de_janeiro <- as_tibble(data_list[[7]])
wmo_seoul <- as_tibble(data_list[[8]])
```

Take a look at the objects.
Do they share any common columns we could use to join them?

## Joining

```{r single-left-join}
# write some code here to join the first two datasets using a left_join()
# then add the assign operator to create a new object called "all_data" or similar
# write a code pipeline here to join all the datasets together
```

## Repeated joins

We want to join all the datasets together, not just two.

```{r join-pipeline}
# write a code pipeline here to join all the datasets together
```

## Plotting wide data

Let's try plotting some of these monthly temperature time series

```{r ggplot-wide}
## try writing code to plot two or more of the cities
```

You should notice that it's not easy to visualize the temperatures of multiple cities Think about why that is.

## Pivot longer

```{r pivot-longer}
## write some code to pivot the dataset long, and save it as a new object "data_long"
```

## Plotting long data

Now try plotting all the cities at once.
Is this easier?

```{r}
# write ggplot code here, think about how to distinguish the different cities with aesthetic mappings
```
